\section{Conclusion}%
\label{sec:conclusion}


\enlargethispage*{3pt}


This paper presented KIF, a framework that uses Wikidata to virtually integrate heterogeneous knowledge sources.
%%
Having Wikidata as the integration model brings some advantages, including a flexible data model and access to a huge vocabulary covering a wide range of topics.
%%
This is particularly important when the integration problem is open-ended: when we don't know which kind of knowledge source we might need to integrate next, or when the sources deal with completely different subjects.
%%
In the domain of chemistry, as illustrated in Section~\ref{sec:background}, when talking about toxicity we often need to refer not only to chemicals but also to things like symptoms, diseases, and even animals---all of these are already covered by Wikidata's vocabulary.
%%
That said, KIF is not restricted to Wikidata's vocabulary.
%%
If needed, the stores can extend their vocabulary with new entities and properties (cf.~the use of PubChem URLs as item ids by the PubChem RDF mapping in Section~\ref{sub:sparql-mapping-for-pubchem}).


% Regarding the architecture of KIF, we tried to conciliate ease of use with flexibility and adherence to standards (SPARQL and RDF).
% %%
% The latter helps to promote the reuse of existing solutions.
% %%
% For example, we currently use Ontop~\cite{Xiao-G-2020} and Comunica~\cite{Taelman-R-2018} in the KIF Middleware, respectively, to construct SPARQL endpoints over SQL endpoints and to aggregate multiple SPARQL endpoints into a single endpoint.
% %%
% We are also studying the integration of other similar solutions: HeFQUIN~\cite{Cheng-S-2024}, as a SPARQL-rewriting-enabled alternative to Comunica, and Squerall~\cite{Mami-M-N-2019}, as a query federation-enabled alternative to Ontop.


Moving to the implementation of KIF, we are currently adding support for parallel requests to the library.
%%
This will speed up not only the mixer but also any store that needs to split large queries into multiple requests.
%%
As we mentioned in Section~\ref{sub:other-store-types-and-methods}, we are also working on an alternative, non-RDF-based implementation of the CSV store.
%%
A more important change is the addition of a mutable store API, which will allow users write onto stores.
%%
The idea is to use SPARQL update queries in the case of the SPARQL store.


Finally, on the theoretical side we are working on the formalization of the semantics of KIF patterns, inspired by the work on WShEx~\cite{Gayo-J-E-L-2022}.


% LocalWords:  KIF PubChem CSV WShEx



%%% Local Variables:
%%% mode: latex
%%% TeX-engine: xetex
%%% TeX-master: "main"
%%% eval: (visual-line-mode 1)
%%% End:
