BW, AC, and BC jointly led the project and contributed to all parts of it. \\
BW worked on all aspects of the project and contributed to all major decisions. He led model design, model training, implemented the majority of the model architecture, and assisted with data selection, elevations, and paper writing.\\
AC co-initiated the project and worked on all aspects of it, including project coordination. Notably, he contributed to monitoring training runs and co-led ablations, final evaluations and paper writing.\\
BC initiated the project and worked on all aspects of it. He contributed to model design and co-led final evaluations, led paper writing, and contributed to the context extension data processing.\\
OW led and conducted the majority of the data selection, processing, and discussion, for all stages of training. He also contributed valuable inputs throughout all stages of the project.\\
OH and ST contributed to a majority of the stages of the project, in particular model architecture and training, with both discussions, implementations and paper writing. Other contributions include pretraining monitoring, final traditional evaluations, and ablations. ST specifically worked on adapting the RoPE kernel for unpadded sequences and running the final GLUE benchmarks. OH additionally conducted a thorough investigation into complex issues that arose during training.\\
RB contributed greatly to the initial evaluation work, focusing on ablations and in-training evals.\\
AG and FL contributed to training efficiency, especially in implementing sequence packing.\\
AG and GA contributed to model evaluations, especially in long context evaluations.\\
TA contributed to discussions throughout the project and assisted in integrating the original research implementation with open source software.\\
NC contributed to context extension data mixtures, and provided insight into model training and on improving the quality of code data.\\
IP and JH provided guidance and support throughout the project, especially on key decisions.