% CVPR 2022 Paper Template
% based on the CVPR template provided by Ming-Ming Cheng (https://github.com/MCG-NKU/CVPR_Template)
% modified and extended by Stefan Roth (stefan.roth@NOSPAMtu-darmstadt.de)

\documentclass[10pt,twocolumn,letterpaper]{article}

%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
\usepackage{cvpr}      % To produce the REVIEW version
%\usepackage{cvpr}              % To produce the CAMERA-READY version
%\usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version

% Include other packages here, before hyperref.
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{ulem}

% It is strongly recommended to use hyperref, especially for the review version.
% hyperref with option pagebackref eases the reviewers' job.
% Please disable hyperref *only* if you encounter grave issues, e.g. with the
% file validation for the camera-ready version.
%
% If you comment hyperref and then uncomment it, you should delete
% ReviewTempalte.aux before re-running LaTeX.
% (Or just hit 'q' on the first LaTeX run, let it finish, and you
%  should be clear).
\usepackage[pagebackref,breaklinks,colorlinks]{hyperref}


% Support for easy cross-referencing
\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}


%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{11345} % *** Enter the CVPR Paper ID here
\def\confName{CVPR}
\def\confYear{2025}
\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Hallo3: Highly Dynamic and Realistic Portrait Image Animation with Diffusion Transformer Networks}

\author{Jiahao Cui\textsuperscript{1},
        Hui Li\textsuperscript{1},
        Yun Zhan\textsuperscript{1},
        Hanlin Shang\textsuperscript{1},
        Kaihui Cheng\textsuperscript{1},
        Yuqi Ma\textsuperscript{1},
        Shan Mu\textsuperscript{1},
        Hang Zhou\textsuperscript{2},\\
        Jingdong Wang\textsuperscript{2},\;\;\;
        Siyu Zhu\textsuperscript{1,3} \\
\textsuperscript{1}Fudan University, \;\;\textsuperscript{2}Baidu Inc,
\;\;\textsuperscript{3}Shanghai Academy of AI for Science
 \\
% {\tt\small xxx@xxx}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
}
\maketitle

%%%%%%%%% ABSTRACT
\begin{abstract}
Existing methodologies for animating portrait images face significant challenges, particularly in handling non-frontal perspectives, rendering dynamic objects around the portrait, and generating immersive, realistic backgrounds.
In this paper, we introduce the first application of a pretrained transformer-based video generative model that demonstrates strong generalization capabilities and generates highly dynamic, realistic videos for portrait animation, effectively addressing these challenges.
The adoption of a new video backbone model makes previous U-Net-based methods for identity maintenance, audio conditioning, and video extrapolation inapplicable. To address this limitation, we design an identity reference network consisting of a causal 3D VAE combined with a stacked series of transformer layers, ensuring consistent facial identity across video sequences.
Additionally, we investigate various speech audio conditioning and motion frame mechanisms to enable the generation of continuous video driven by speech audio. Our method is validated through experiments on benchmark and newly proposed wild datasets, demonstrating substantial improvements over prior methods in generating realistic portraits characterized by diverse orientations within dynamic and immersive scenes. Further visualizations and the source code are available at: \href{https://fudan-generative-vision.github.io/hallo3/#/}{https://fudan-generative-vision.github.io/hallo3}.

\end{abstract}

%%%%%%%%% BODY TEXT
\input{sections/arxiv_introduction}
\input{sections/related_work}
%\input{sections/preliminary}
\input{sections/arxiv_method}
\input{sections/arxiv_experiments}
\input{sections/arxiv_conclusion}

%%%%%%%%% REFERENCES
{\small
\bibliographystyle{ieee_fullname}
\bibliography{egbib}
}
% \input{sections/appendix}

\end{document}