\section{3D Gaussian Splatting and Surface Reconstruction}

Our method relies on the original 3D Gaussian Splatting~(3DGS) method~\cite{kerbl3Dgaussians} for initialization and on SuGaR~\cite{guedon2023sugar} to align Gaussians with the surface of the scene and facilitate the extraction of a mesh. We briefly describe 3DGS and SuGaR in this section before describing our method in the next section.


\subsection{3D Gaussian Splatting} 

3DGS represents the scene as a large set of Gaussians. Each Gaussian $g$ is equipped with a mean $\mu_g\in \IR^3$ and a positive-definite covariance matrix $\Sigma_g\in \IR^{3\times 3}$. The covariance matrix is parameterized by a scaling vector $s_g\in\IR^3$ and a quaternion $q_g\in\IR^4$ encoding the rotation of the Gaussian. 

In addition, each Gaussian has a view-dependent radiance represented by an opacity $\alpha_g\in [0,1]$ and a set of spherical harmonics coordinates defining the colors emitted for all directions. To render an image from a given viewpoint, a rasterizer ``splats'' the 3D Gaussians into 2D Gaussians parallel to the image plane and blends the splats depending on their opacity and depth. This rendering is extremely fast, which is one of the advantages of 3DGS over volumetric rendering as in NeRFs for example~\cite{mildenhall2020nerf, mueller2022instantngp, barron2022mipnerf360}.


Gaussian Splatting can be seen as an approximation of the traditional volumetric rendering of radiance fields with the following density function $d$, computed as the sum of the Gaussian values weighted by their alpha-blending coefficients at any 3D point $p\in \IR^3$:
%
 \begin{equation}
    d(p) = \sum_{g} \alpha_g \exp\left(-\frac{1}{2}(p - \mu_g)^T \Sigma^{-1}_g (p - \mu_g)\right) \> .
    \label{eq:gaussian_splatting_density}
\end{equation}

We initialize our Gaussian Frosting method using a vanilla 3DGS optimization: Gaussians are initialized using the point cloud produced by an SfM~\cite{snavely-2006-structure-from-motion} algorithm like COLMAP~\cite{schoenberger2016mvs,schoenberger2016sfm}, required to compute camera poses. The Gaussians' parameters (3D means, scaling vectors, quaternions, opacities, and spherical harmonics coordinates) are then optimized to make the renderings match the ground truth images of the scene, using a rendering loss that only consists in a combination of a pixel-wise L1 distance and a more structural D-SSIM term. 


\subsection{SuGaR Mesh Extraction} 

Vanilla 3DGS does not have regularization explicitly encouraging Gaussians to align with the true surface of the scene. Our Gaussian Frosting representation relies on a mesh that approximates this surface, in order to be editable by traditional tools. To obtain this mesh, we rely on the method proposed in SuGaR~\cite{guedon2023sugar}, which we improve by automatically selecting a critical hyperparameter.

SuGaR proposes a regularization term encouraging the alignment of the 3D Gaussians with the true surface of the scene during the optimization of Gaussian Splatting, as well as a mesh extraction method. After enforcing the regularization, the optimization provides Gaussians that are mostly aligned with the surface albeit not perfectly: We noticed that in practice, a large discrepancy between the regularized Gaussians and the extracted mesh indicates the presence of fuzzy materials or surfaces that require volumetric rendering. We thus exploit this discrepancy as a cue to evaluate where the Frosting should be thicker.

\input{figures/pipeline}

\section{Creating a Frosting Layer from Images}

In this section, we describe our Gaussian Frosting creation method: 
First, we extract an editable surface with optimal resolution using SuGaR. We then detail how we use this surface-based model to go back to a volumetric but editable representation built around the mesh. This representation adapts to the complexity of the scene and its need for more volumetric effects. Finally, we describe how we parameterize and refine this representation. An overview is provided Figure~\ref{fig:frosting-pipeline}.

\input{4_1}

\input{4_2_backward}

\subsection{Frosting Optimization}

Once we constructed the outer and inner bounds of the Frosting layer, we initialize a densified set of Gaussians inside this layer and optimize them using 3DGS rendering loss as the unconstrained Gaussians. To make sure the Gaussians stay inside the frosting layer during optimization, we introduce a new parameterization of the Gaussians. Moreover, this parameterization will make possible to 
easily adjust the Gaussians' parameters when editing the scene.


\subsubsection{Parameterization.} Let us consider a triangular face of the base mesh $\calM$, with vertices denoted by~$\vec{v}_0, \vec{v}_1$, and~$\vec{v}_2$ and their corresponding normals~$\vec{n}_0, \vec{n}_1$, and~$\vec{n}_2$. After extracting inner and outer shifts from unconstrained Gaussians, we obtain six new vertices $(\vec{v}_i+\innershift_i \vec{n}_i)_{i=0,1,2}$ and $(\vec{v}_i+\outershift_i \vec{n}_i)_{i=0,1,2}$ that respectively belong to the inner and outer bounds of the frosting.
%
Specifically, these six vertices delimit an irregular triangular prism. We will refer to such polyhedrons as ``prismatic cells''. 
%
We parameterize the 3D mean~$\mu_g\in \IR^3$ of a Gaussian~$g \in \calG$ located inside a prismatic cell with a set of six barycentric coordinates split into two subsets~$(b_g^{(i)})_{i=0,1,2}$ and $(\beta_g^{(i)})_{i=0,1,2}$, such that 
%
\begin{equation}
    \mu_g = \sum_{i=0}^2 \left(
    b_g^{(i)} \left(\vec{v}_i+\outershift_i \vec{n}_i\right) +
    \beta_g^{(i)} \left(\vec{v}_i+\innershift_i \vec{n}_i\right)\right) \> ,
    \label{eq:barycentric-coordinates}
\end{equation}
%
with barycentric coordinates verifying $\sum_{i=0}^2 ( b_g^{(i)}+\beta_g^{(i)}) = 1$.
%
Using barycentric coordinates enforces Gaussians to stay inside their corresponding prismatic cell, and guarantees the stability of our representation during optimization.
%
In practice, we apply a softmax activation on the parameters to optimize to obtain barycentric coordinates that sum up to 1. 

\subsubsection{Initialization.} For a given budget $N$ of Gaussians provided by the user, we initialize $N$ Gaussians in the scene by sampling $N$ 3D centers $\mu_g$ in the frosting layer. Specifically, for sampling a single Gaussian, we first randomly select a prismatic cell with a probability proportional to its volume. Then, we sample random coordinates that sum up to 1. 
This sampling allows for allocating more Gaussians in areas with fuzzy and complex geometry, where more volumetric rendering is needed. However, flat parts  in the layer may also need a large number of Gaussians to recover texture details. Therefore, in practice, we instantiate $N/2$ Gaussians with uniform probabilities in the prismatic cells, and $N/2$ Gaussians with probabilities proportional to the volume of the cell.

We initialize the colors of the Gaussians with the color of the closest Gaussian in the unconstrained representation. However, we do not use the unconstrained Gaussians to initialize opacity, rotation, and scaling factors, as in practice, following the strategy from 3DGS~\cite{kerbl3Dgaussians} for these parameters provides better performance: 
We suppose the positions and configuration of the Gaussians inside the Frosting layer are already a good initialization, and resetting opacities, scaling factors and rotations helps Gaussians to take a fresh start, avoiding a potential local minimum encountered by previous unconstrained Gaussians.

Our representation allows for a much better control over the number of Gaussians than the original Gaussian Splatting densification process, as it is up to the user to decide on a number of Gaussians to instantiate in the frosting layer. These Gaussians will be spread in the entire frosting in a very efficient way, adapting to the need for volumetric rendering in the entire scene.

\subsubsection{Optimizing the Gaussian Frosting.} We reload the unconstrained Gaussians and apply our method for computing the inner and outer bounds of the Frosting. Then, for a given budget of $N$ Gaussians, we initialize $N$ Gaussians in the Frosting and optimize the representation while keeping the number of Gaussians constant. Note that compared to Vanilla 3DGS, this allows to control precisely the number of Gaussians.


\subsubsection{Editing, Deforming, and Animating the Frosting.} When deforming the base mesh, the positions of Gaussians automatically adjust in the frosting layer thanks to the use of the barycentric coordinates. 
%
To automatically adjust the rotation and scaling factors of the Gaussians, we propose a strategy different from the surface-based adjustment from SuGaR: In a given prismatic cell with center $\vec{c}$ and vertices $\vec{v}_i$ for $0\leq i<5$, we first estimate the local transformation at each vertex $\vec{v}_i$ by computing the rotation and rescaling of the vector $(c - \vec{v}_i)$. 
%
Then, we use the barycentric coordinates of a Gaussian $g$ to compute an average transformation at point $\mu_g$ from the transformation of all 6 vertices, and we adjust the rotation and scaling factors of $g$ by applying this average transformation. 
%
Please note that the spherical harmonics are also adjusted in practice, to ensure the consistency of the emitted radiance depending on the averaged rotation applied to the Gaussian.
%
We provide more details about this automatic adjustment of Gaussian parameters in the supplementary material.