\newpage
\centerline{\large Supplementary Material}

\vspace{0.3in}
\noindent
In this supplementary material, we provide the following elements:
\begin{itemize}
    \item A description of our method to improve the surface reconstruction from SuGaR~\cite{guedon2023sugar}.
    \item Additional details about our strategy to initialize the Frosting layer and automatically adjust Gaussians' parameters when deforming, editing, or animating our representation.
\end{itemize}
% TODO
We also provide a \href{https://anttwo.github.io/frosting/}{\underline{video}} that offers an overview of the approach and showcases additional qualitative results. Specifically, the video demonstrates how Frosting can be used to edit, combine or animate Gaussian Splatting representations.

\setcounter{section}{+6}
\section{Improving surface reconstruction}

\input{4_1_improving_surf_rec}

\section{Initializing the frosting layer}

\subsection{Sampling Gaussians in the frosting layer} 

\subsubsection{Sampling more Gaussians in thicker parts of the frosting.} For a given budget $N$ of Gaussians provided by the user, we initialize $N$ Gaussians in the scene by sampling $N$ 3D centers $\mu_g$ in the frosting layer. Specifically, for sampling a single Gaussian, we first randomly select a prismatic cell with a probability proportional to its volume. Then, we sample random coordinates that sum up to 1. 
%
This sampling allows for allocating more Gaussians in areas with fuzzy and complex geometry, where more volumetric rendering is needed. However, flat parts  in the layer may also need a large number of Gaussians to recover texture details. Therefore, in practice, we instantiate $N/2$ Gaussians with uniform probabilities in the prismatic cells, and $N/2$ Gaussians with probabilities proportional to the volume of the cell.

\subsubsection{Contracting volumes in unbounded scenes.} In real unbounded scenes, 3D Gaussians located far away from the center of the scene can have a significantly large volume despite their limited participation in the final rendering. This can lead to an unnecessarily large number of Gaussians being sampled in the frosting layer far away from the training camera poses. To address this issue, we propose distributing distant Gaussians proportionally to disparity (inverse distance) rather than distance.

When sampling Gaussians in practice, we start by contracting the volumes of the prismatic cells. We achieve this by applying a continuous transformation $f:\IR^3 \rightarrow \IR^3$ to the vertices of the outer and inner bounds of the frosting layer. 
%
Then, we compute the volumes of the resulting ``contracted'' prismatic cells and use these adjusted volumes for sampling Gaussians within the frosting layer, as previously described. The transformation function $f$ aims to contract the volume of prismatic cells located far away from the center of the scene. We define $f$ using a formula similar to the contraction transformation introduced in Mip-NeRF~360~\cite{barron2022mipnerf360}:
\begin{equation}
    f(x) = \begin{cases}
        x & \text{if} \>\>\>\> \|x-c\| \leq l\\
        c + l \times \left(2 - \frac{l}{\|x-c\|}\right) \left(\frac{x-c}{\|x-c\|}\right) & \text{if} \>\>\>\> \|x-c\| > l
    \end{cases} \>,
    \label{eq:contraction}
\end{equation}
where $c\in \IR^3$ is the center of the bounding box containing all training camera positions, and $l\in\IR_+$ is equal to half the length of the diagional of the same bounding box. 
%
We choose the bounding box of the camera positions as our reference scale because both 3D Gaussian Splatting~\cite{kerbl3Dgaussians} and SuGaR~\cite{guedon2023sugar} use this same reference for scaling learning rates and distinguishing foreground from background in unbounded scenes.

\subsection{Avoiding self-intersections in the frosting layer}

In the main paper, we define the inner and outer bounds of the frosting layer by adding inner and outer shifts $\innershift_i$ and $\outershift_i$ to the vertices $\vec{v}_i$ of the base mesh. This results in two bounding surfaces with vertices $\vec{v}_i + \innershift_i$ and $\vec{v}_i + \outershift_i$. 
%
In practice, we wish to minimize self-intersections within the frosting layer, specifically avoiding prismatic cells intersecting with each other.

While self-intersections do not directly impact rendering quality, they can lead to artifacts during scene editing or animation. Consider the scenario where different cells intersect. In such cases, moving a specific triangle of the base mesh may not affect all Gaussians intersecting the surrounding cell: Some Gaussians may belong to prismatic cells associated with different triangles, resulting in artifacts due to their failure to follow local motion or scene edits.

To mitigate self-intersections, we adopt an indirect approach for initializing the shifts $\innershift$ and $\outershift$. 
Instead of using the final computed values directly, we start with shifts equal to zero and progressively increase them until reaching their final values. 
As soon as an inner vertex (or outer vertex) of a prismatic cell is detected to intersect another cell, we stop further increases in its inner shift (or outer shift). 
This straightforward process significantly reduces self-intersections in the frosting layer while maintaining rendering performance.

By following this approach, we ensure that the frosting layer remains free from unwanted artifacts while preserving efficient rendering capabilities.

\section{Adjusting Gaussians' parameters for edition}

When editing or animating the scene, we automatically adjust Gaussians' parameters. Specifically, in a given prismatic cell with center $\vec{c}$ and six vertices $\vec{v}_i$ for $0\leq i<6$, we first estimate the local transformation at each vertex $\vec{v}_i$ by computing the rotation and rescaling of the vector $(c - \vec{v}_i)$. 

To compute the local rotations at vertex $\vec{v}_i$, we use an axis-angle representation where the axis angle is the normalized cross-product between the previous and the current values of the vector $(c - \vec{v}_i)$.
%
The local rescaling transformation at vertex $\vec{v}_i$ is computed as the transformation that scales along axis $(c - \vec{v}_i)$ with the appropriate factor but leaves other axes unchanged.

To update the scaling factors and rotation of a Gaussian $g$, we first apply each of these six transformations on the three main axes of the Gaussian. We then average the resulting axes using the barycentric coordinates of the center $\mu_g$ of the Gaussian. We finally orthonormalize the three resulting axes.
