\section*{Appendix}\label{sec:appendix}

\subsection{Dataset}

%% JL: This is already mentioned in the text.
%% Please emove the table
\input{tables/rw_corpus_features}

\subsection{Metadata Content Example}
\label{ssec:appendix.metadataexample}
% \begin{verbatim}
% {
%     "meta": {
%         "docid": "wikipedia/en/12/56",
%         "url": "https://en.wikipedia.org/wiki/LLM",
%         "title": "LLM",      
%         "download_date": "2024-04-01",
%         "language": "en",
%         "language_score": 0.97
%         },
%     "text": "LLM may refer to:\n
% - Master of Laws (Latin: Legum Magister), a postgraduate degree\n 
% - LLM Communications, a defunct lobbying firm \n
% - LLM Lettering, a typeface\n
% - Large language model, the use of large neural networks for language modeling\n
% - Logic learning machine, a machine learning method\n"
% }
% \end{verbatim}

\begin{verbatim}
    {
    "meta": {
        "url": "https://en.wikipedia.org/wiki/Organic%20Chemistry/Cover", 
        "title": "Organic Chemistry/Cover", 
        "docid": "wikimedia/wikibooks/enwikibooks-20240401", 
        "language": "en", 
        "language_score": 1.0, 
        "download_data": "2024-04-01", 
        }
    "text": "Welcome to the world's foremost open contentOrganic
            Chemistry Textbookon the web!\n\n
            The Study of Organic Chemistry[...]", 
    }
\end{verbatim}


\subsection{Curated Data}
\label{sec:appendix.curated}

\subsubsection{Filtering}
\label{sec:appendix.curated.filtering}
%%JL why did we look at correlation?
The correlation between the percentage filtered and the total number of words in a dataset was relatively low at \( r = 0.26 \) (p = 0.022), suggesting only a weak relationship. Similarly, a slightly stronger correlation was observed between the percentage filtered and the number of documents, at \( r = 0.33 \) (p = 0.003), indicating that datasets with more documents tend to require somewhat more filtering, but the relationship is still far from definitive.

We also examined the impact of format on filtering, but found no significant correlation (\( r = 0.03 \), p = 0.79). However, when considering the language of the dataset, we found a more notable correlation (\( r = 0.37 \), p = 0.001), hinting that certain languages may be more prone to needing extensive filtering, possibly due to differences in data quality or availability across languages.

Finally, domain showed no significant correlation with the percentage of data filtered (\( r = -0.004 \), p = 0.97).

These results suggest that filtering operates somewhat independently from these broader dataset characteristics. Language differences, in particular, seem to warrant more focused attention in future analyses, as they exhibit a stronger connection to filtering needs compared to other factors. 

\subsubsection{List of Curated Datasets}
The complete list of curated dataset is presented in Table \ref{tab:curated_data_list} (next page).


\input{tables/curated_datasets_nicolo}
%\input{tables/curated_datasets_johannes}

% \newpage
% \include{tables/global\_langauges}
