\begin{tcolorbox}[title = {Case 1}, colframe=darkgreen, colback=lightgray, breakable]
\small
\textbf{Text:}\\Photodynamic diagnosis (PDD) is available at some centres. PDD makes cancer cells visible under violet light. This makes abnormal growths (tumours) easier to find and remove. More complete removal reduces the risk of cancer coming back.PDD is used in urology to test for bladder cancer. A catheter is inserted through the urethra and into the bladder. A solution of 5-aminolaevulinic acid or hexaminolaevulinic acid is passed through the catheter into the bladder. Then the catheter is taken out immediately. Cancer cells in the bladder process the chemicals in the solution and become fluorescent under violet light. No side-effects or complications have been reported for PDD.\\
\textbf{Rating:}
\\
\begin{tabular}{@{}llll}
Educational Value: & 99 & Expertise: & 92 \\ 
Fact and Trivia: & 92 & Scarcity: & 85 \\
Reasoning Level: & 93   & Story-likenss: & 55 \\ 
Structural format: & 91 & Subjectivity: & 25 \\
\end{tabular}
\\
\textbf{Tagging:}
\\
\begin{tabular}{@{}ll}
First Level Tag: & Medical And Health \\
Second Level Tag: & Medical Procedures \\
Third Level Tag: & Diagnostic Procedures \\
\end{tabular}
\\
\textbf{Editing:}
\\
Photodynamic diagnosis (PDD) is accessible at certain facilities, a technique that enables the visualization of cancer cells under violet light, thereby enhancing the detection and removal of irregularities, such as tumours, with increased precision. This method further minimizes the likelihood of cancer cells recurring. In the field of urology, PDD is utilized for the identification of bladder cancer. The procedure involves the insertion of a catheter through the urethra into the bladder. Within the catheter, a solution containing 5-aminolaevulinic acid or hexaminolaevulinic acid is delivered. Following this, the catheter is promptly removed. It is observed that cancer cells within the bladder respond to these chemicals, converting them into fluorescent particles under the influence of violet light. Due to its safe and effective operation, no adverse effects or issues have been noted associated with PDD.
\end{tcolorbox}

\begin{tcolorbox}[title = {Case 2}, colframe=darkgreen, colback=lightgray, breakable]
\small
\textbf{Text:}\\Tron (TRX) is an ambitious project which aims to own the bragging rights regarding the entertainment industry. To make that a reality, a day hardly ends...Tron (TRX) is undoubtedly an ambitious project in the crypto community, and it aims to own the bragging rights of the entertainment industry. However, they are...Before I start explaining how Tron (TRX) Mainnet, Odyssey, is enhancing the chances of Tron adoption already, I need to explain some crucial thoughts with you....What will be the price of Tron (TRX) in five years from now? Tron is one of the top cryptocurrencies in the world (always in the spotlight). The cryptocurrency’s mainnet is now live, and on its independence day, it...Pornhub, one of the largest X-rated video sites, expands its virtual currency payment options announcing they will be soon accepting Tron (TRX). In recent times, Tron...Can Tron (TRX) become one of the top 5 cryptos in the market now?In recent days, Tron (TRX) has been experiencing one of the biggest tractions of the market, a fact that without a doubt is helping the token...\\
\textbf{Rating:}
\\
\begin{tabular}{@{}llll}
Educational Value: & 45 & Expertise: & 54 \\ 
Fact and Trivia: & 80 & Scarcity: & 10 \\
Reasoning Level: & 51   & Story-likenss: & 60 \\ 
Structural format: & 55 & Subjectivity: & 80 \\
\end{tabular}
\\
\textbf{Tagging:}
\\
\begin{tabular}{@{}ll}
First Level Tag: & Technology and Internet \\
Second Level Tag: & Entertainment and Media \\
Third Level Tag: & Digital Media \\
\end{tabular}
\\
\textbf{Editing:}
\\
Tron (TRX) embodies a visionary endeavor to dominate the boundaries of the entertainment sector, a goal that promises to be achieved on a nearly perpetual basis. In pursuit of this ambitious objective, Tron (TRX) continually strives to enhance its presence. In the realm of cryptocurrency, Tron (TRX) stands out for its bold ambitions, vying for control over the entertainment industry's prestige. Despite its aspirations, Tron (TRX) is already making strides in securing a larger user base, thanks to its innovative Mainnet, Odyssey. This initiative is pivotal in laying the groundwork for Tron's broader adoption. As we reflect on the potential future value of Tron (TRX) in the coming five years, it is noteworthy that Tron holds a prominent position among the world's cryptocurrencies, drawing significant attention. The cryptocurrency's flagship network has commenced operations, and its founding day marks an important milestone for its development. The entertainment sector's leading platform, Pornhub, announces a expansion into virtual currency payments, revealing plans to begin accepting Tron (TRX) as a payment method shortly. This move highlights the growing acceptance and adoption of Tron (TRX) within the digital landscape. Furthermore, there exists a discussion surrounding the potential for Tron (TRX) to surpass other top 5 cryptos in the market, with recent market trends clearly indicating its robust growth and widespread acceptance.
\end{tcolorbox}

\begin{tcolorbox}[title = {Case 3}, colframe=darkgreen, colback=lightgray, breakable]
\small
\textbf{Text:}
\\
Gown, \$12,900, Valentino, Bal Harbour Shops and Design District; diamond stud earrings, price upon request, elanjewels.us.
Jennifer Hudson, emotion is everything. It’s how she breathes life into a character. It’s how she makes a song explode. And it’s why—since Hudson was a child—people are drawn to her talent like a moth to a flame. Well, wait until you see her newest film.
Larkin coat, \$6,770, by Erdem at Saks Fifth Avenue, Bal Harbour Shops, Brickell City Centre and Dadeland Mall; satin Bullet bodysuit, \$350, by Fleur du Mal at Intermix, Bal Harbour Shops, Brickell City Centre and Lincoln Road; Kimmy belt, \$625, at Isabel Marant, Design District; printed velvet trousers, \$900, by Paco Rabanne at The Webster, Bal Harbour Shops and South Beach; Ellabrita strass sandal 105, \$1,150, by René Caovilla at Neiman Marcus, Bal Harbour Shops and Shops at Merrick Park; diamond earrings, price upon request, at elanjewels.us.
Gown, \$25,000, Valentino, Bal Harbour Shops and Design District; feather boa, \$3,990, Loewe, Design District.
\\
\textbf{Rating:}
\\
\begin{tabular}{@{}llll}
Educational Value: & 10 & Expertise: & 2 \\ 
Fact and Trivia: & 51 & Scarcity: & 1 \\
Reasoning Level: & 11   & Story-likenss: & 50 \\ 
Structural format: & 36 & Subjectivity: & 63 \\
\end{tabular}
\\
\textbf{Tagging:}
\\
\begin{tabular}{@{}ll}
First Level Tag: & Fashion and Beauty \\
Second Level Tag: & Fashion Apparel \\
Third Level Tag: & Formal Attire \\
\end{tabular}
\\
\textbf{Editing:}
\\
A gown from Valentino, Bal Harbour Shops and Design District, retails for \$12,900. Additionally, diamond stud earrings are available at a custom price, with the source being elanjewels.us. Jennifer Hudson is celebrated for her profound emotional depth, capturing the essence of her characters and elevating musical pieces to new heights. Her innate talent, which has attracted a multitude of fans since she was a young age, is poignantly depicted in her latest cinematic venture.
Erdem's Larkin coat, priced at \$6,770, is showcased at Saks Fifth Avenue, alongside other collections in Bal Harbour Shops, Brickell City Centre, and Dadeland Mall. The Bullet bodysuit, priced at \$350, features a satin material by Fleur du Mal at Intermix, along with other merchandise in these same locations. A belt by Kimmy, priced at \$625, is available at Isabel Marant in the Design District. For a more contemporary look, a printed velvet trouser, priced at \$900, by Paco Rabanne is offered at The Webster in Bal Harbour Shops and South Beach.
Elenabrita's Ellabrita strass sandal 105, priced at \$1,150, is designed by René Caovilla and available at Neiman Marcus, Shops at Merrick Park, and additional retailers. Diamond earrings, once requested, can be purchased from elanjewels.us.
A gown from Valentino, priced at \$25,000, is available from Bal Harbour Shops and Design District, while a feather boa, priced at \$3,990, adds a distinctive touch to Loewe's designs in the Design District.
\end{tcolorbox}