\begin{python}  
system_prompt = f"You are an expert question-answer generation based on the given image. 
    You takes a image as input and generate question and answer pairs about the image as output."
user_prompt = f"Your task is to generate {number} question and its' respective answer pairs based on the given image. Generate question about object, size, shape, color or overall context and also generate answer for each questions. 
    Respond with each question in between <QUESTION> and </QUESTION> and respond with each concise answer in between <ANSWER> and </ANSWER>, eg: 
    1. <QUESTION>Question 1</QUESTION>\n 
    2. <QUESTION>Question 2</QUESTION>\n 
    3. <QUESTION>Question 3</QUESTION>\n"
    if n>3:
        for i in range(4, n):
            user_prompt += f"{n}. <QUESTION>Question {n}</QUESTION>\n"
\end{python}
To utilize LLM as a QA response, we employ following template.
\begin{python}
system_prompt = f"You are an expert answer agent based on the given text. \
    You takes a text as input and respond correct answer as output."
user_prompt = f"Your task is to respond on following three questions based on the context. Respond with answers in between <ANSWER> and </ANSWER>, eg: \
    1. <ANSWER>ANSWER 1</ANSWER>\n \
    2. <ANSWER>ANSWER 2</ANSWER>\n \
    3. <ANSWER>ANSWER 3</ANSWER>\n \
    Context: {described_text}\n"
    for n in range(len(Q_list)):
        user_prompt += f'Question 1: {Q_list[n]}\n'
\end{python}
To utilize LLM as evaluator of responses, we employ following template.
\begin{python}
system_prompt = f"You are an expert evaluator agent based on the given question and answer. \
    You takes a respond as input and evaluate whether it is correct or not as output."
user_prompt = f""
score = 0
for n in range(len(A_list)):
    user_prompt = f"Question is \"{Q_list[n]}\" and the correct answer is {A_list[n]}.\n"
    user_prompt += f"My respond: {respond_list[n]}\n"
    user_prompt += f"Your task is to evaluate my respond based on the question and correct answer. Write <CORRECT> if it is correct, write <WRONG> if it is incorrect. And provide the reason of your evaluation.\n"
\end{python}