\documentclass[]{fairmeta}
% Option "twocolumn" available, but please prioritize single-column

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{colortbl}
\usepackage{overpic}
\usepackage{amsthm}

\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{scalerel}
\newcommand{\sicong}[1]{{\color{blue}#1}}
\usepackage{makecell}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}

% define ours package
\usepackage{nicefrac}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{bbding}
\usepackage{pifont}
\usepackage{color, colortbl}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}} % \tabincell{c}{
% \newcommand{\hr}[1]{\textcolor{blue}{#1}}
\newcommand{\hr}[1]{\textcolor{black}{#1}}
\definecolor{gray}{gray}{0.9}
\newcommand{\ssymbol}[1]{^{\@fnsymbol{#1}}}
\def\R{\mathbb{R}}
\newcommand{\Lleft}{\left(}
\newcommand{\Rright}{\right)}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}

\newcommand{\modelname}{LongVU}

\title{Efficient Track Anything}

\author[1,\dagger]{Yunyang Xiong}
\author[1,2]{Chong Zhou}
\author[1]{Xiaoyu Xiang}
\author[1]{Lemeng Wu}
\author[1]{Chenchen Zhu}
\author[1]{Zechun Liu}
\author[1]{Saksham Suri}
\author[1]{Balakrishnan Varadarajan}
\author[1]{Ramya Akula}
\author[1]{Forrest Iandola}
\author[1,\dagger]{Raghuraman Krishnamoorthi}
\author[1,\dagger]{Bilge Soran}
\author[1,\dagger]{Vikas Chandra}

\affiliation[1]{Meta AI}
\affiliation[2]{Nanyang Technological University}

\contribution[\dagger]{Project lead}

\abstract{Segment Anything Model 2 (SAM 2) has emerged as a powerful tool for video object segmentation and tracking anything. Key components of SAM 2 that drive the impressive video object segmentation performance include a large multistage image encoder for frame feature extraction and a memory mechanism that stores memory contexts from past frames to help current frame segmentation. The high computation complexity of multistage image encoder and memory module has limited its applications in real-world tasks, e.g., video object segmentation on mobile devices. To address this limitation, we propose EfficientTAMs, lightweight track anything models that produce high-quality results with low latency and model size. Our idea is based on revisiting the plain, nonhierarchical Vision Transformer (ViT) as an image encoder for video object segmentation, and introducing an efficient memory module, which reduces the complexity for both frame feature extraction and memory computation for current frame segmentation. We take vanilla lightweight ViTs and efficient memory module to build EfficientTAMs, and train the models on SA-1B and SA-V datasets for video object segmentation and track anything tasks. We evaluate on multiple video segmentation benchmarks including semi-supervised VOS and promptable video segmentation, and find that our proposed EfficientTAM with vanilla ViT perform comparably to SAM 2 model (HieraB+SAM 2) with $\sim$2x speedup on A100 and $\sim$2.4x  parameter reduction. On segment anything image tasks, our EfficientTAMs also perform favorably over original SAM with $\sim$20x  speedup on A100 and $\sim$20x  parameter reduction. On mobile devices such as iPhone 15 Pro Max, our EfficientTAMs can run at $\sim$10 FPS for performing video object segmentation with reasonable quality, highlighting the capability of small models for on-device video object segmentation applications.}

% \date{\today}
\correspondence{\email{yunyang@meta.com}}

\metadata[Project]{\url{https://yformer.github.io/efficient-track-anything/}}

\begin{document}

\maketitle

\input{sec/1_intro}
\input{sec/2_related}
\input{sec/3_method}
\input{sec/4_experiments}
\input{sec/5_conclusion}
\input{sec/7_acknowledgements}

\bibliographystyle{assets/plainnat}
\bibliography{main}

\clearpage
\newpage
\beginappendix

\input{sec/6_suppl}

\end{document}