\documentclass[11pt]{article}
\pdfoutput=1

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{amssymb}
\usepackage{bm}
\usepackage{bbm}
\usepackage{amsmath}  % Define \boldsymbol (in amsbsy too) and align
\usepackage{amsthm}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage[margin=1in]{geometry}
\usepackage{authblk}
\usepackage[numbers, sort]{natbib}
\usepackage{xspace}
\usepackage{comment}
\usepackage{thmtools, thm-restate}  % for repeating thms in appendix
\usepackage{verbatim}
\usepackage[T1]{fontenc}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}


% \usepackage{hyperref}
\definecolor{tabblue}{HTML}{5555CC}
\usepackage[hidelinks,colorlinks=true,linkcolor=tabblue,citecolor=tabblue]{hyperref}

% \usepackage[hidelinks,colorlinks=true,linkcolor=tabblue,citecolor=tabblue]{hyperref}

% % \usepackage[hidelinks,colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{url}

\usepackage[capitalize,noabbrev]{cleveref}
% \usepackage{subcaption}
\usepackage{caption}
\usepackage{pifont}
\usepackage{makecell}
\usepackage{tabularx}
\usepackage{float}
% % Tables
\usepackage{adjustbox}
\usepackage{varwidth}
%
\input{template/math}
\input{template/style}

\newcommand*{\ShowNotes}{}

\newif\ifarxiv

%
\title{Effectively Modeling Time Series with \\Simple Discrete State Spaces}
%
\author{Michael Zhang$^*$, Khaled Saab\thanks{ Equal Contribution. Order determined by forecasting competition.}\;, Michael Poli,  Tri Dao, Karan Goel, and Christopher R\'{e} \\
% about author (webpage, alternative address)---\emph{not} for acknowledging
% funding agencies.  Funding acknowledgements go at the end of the paper.} \\
Stanford University \\
\vspace{0.25cm}
\texttt{mzhang@cs.stanford.edu}, \texttt{\{ksaab,poli\}@stanford.edu}, \texttt{\{tridao,kgoel,chrismre\}@cs.stanford.edu}
% \AND
% Coauthor \\
% Affiliation \\
% Address \\
% \texttt{email}
}
%
% \iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
\begin{document}
%
\maketitle

%
\doparttoc
\faketableofcontents
%
\input{sections/abstract}
\input{sections/introduction}
\input{sections/preliminaries}
\input{sections/method}
\input{sections/experiments}
\input{sections/conclusion}

\newpage

\section{Ethics Statement}
A main objective of our work is to improve the ability to classify and forecast time series, which has real-world applications in many fields. These applications may have high stakes, such as classifying abnormalities in medical time series. In these situations, incorrect predictions may lead to harmful patient outcomes. It is thus critical to understand that while we aim to improve time series modeling towards these applications, we do not solve these problems. Further analysis and development into where models fail in time series modeling is necessary, including potentials intersections with research directions such as robustness and model biases when aiming to deploy machine learning models in real world applications.  
% 
% In many applications, such as classifying abnormal medical time series, incorrect predictions may lead to harmful patient outcomes. For many such applications, it is critical to carry out a deeper analysis and understanding of model failures, including an investigation into potential model biases, where the model may perform well on average but underperform on certain subpopulations. 
\section{Reproducibility}
We include code for the main results in Table~\ref{tab:informer-s-long} at \href{https://github.com/HazyResearch/spacetime}{https://github.com/HazyResearch/spacetime}. We provide training hyperparameters and dataset details for each benchmark in Appendix~\ref{app:exp_details}, discussing the Informer forecasting benchmark in Appendix~\ref{app:informer_details}, the Monash forecasting benchmark in Appendix~\ref{app:monash_details}, and the ECG and speech audio classification benchmarks in Appendix~\ref{app:classification_exps}.
%
We provide proofs for all propositions and algorithm complexities in  Appendix~\ref{appendix:theory}. 

\section{Acknowledgements}
We thank Albert Gu, Yining Chen, Dan Fu, Ke Alexander Wang, and Rose Wang for helpful discussions and feedback. We also gratefully acknowledge the support of NIH under No. U54EB020405 (Mobilize), NSF under Nos. CCF1763315 (Beyond Sparsity), CCF1563078 (Volume to Velocity), and 1937301 (RTML); US DEVCOM ARL under No. W911NF-21-2-0251 (Interactive Human-AI Teaming); ONR under No. N000141712266 (Unifying Weak Supervision); ONR N00014-20-1-2480: Understanding and Applying Non-Euclidean Geometry in Machine Learning; N000142012275 (NEPTUNE); NXP, Xilinx, LETI-CEA, Intel, IBM, Microsoft, NEC, Toshiba, TSMC, ARM, Hitachi, BASF, Accenture, Ericsson, Qualcomm, Analog Devices, Google Cloud, Salesforce, Total, the HAI-GCP Cloud Credits for Research program,  the Stanford Data Science Initiative (SDSI), and members of the Stanford DAWN project: Facebook, Google, and VMWare. The U.S. Government is authorized to reproduce and distribute reprints for Governmental purposes notwithstanding any copyright notation thereon. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views, policies, or endorsements, either expressed or implied, of NIH, ONR, or the U.S. Government.



%
\bibliographystyle{abbrvnat}
\bibliography{main.bib}
%
\newpage
%
\begin{center}
    \huge\bf{Appendix:\\ Effectively Modeling Time Series with \\Simple Discrete State Spaces}
\end{center}

\appendix
\addcontentsline{toc}{section}{}
\part{}
\parttoc
%
\input{sections/related_work}
\input{sections/appendix/theory}
\input{sections/appendix/02_details}
\input{sections/appendix/experiment_details}
\input{sections/appendix/experiment_results}
%
\end{document}