\documentclass{article} % For LaTeX2e
\usepackage{iclr2021_conference,times}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}
\usepackage{amssymb} 
\usepackage{amsmath} 
\usepackage{amsthm} 
\usepackage{amsfonts} 
\usepackage{enumitem} 
\usepackage{tabularx} 
\usepackage{centernot}
\usepackage[parfill]{parskip}
\usepackage{quiver}

\usepackage{graphicx} 
\usepackage{scalerel}
\newtheorem{theorem}{Theorem} 

\theoremstyle{definition} 
% \newtheorem{problem}{Exercício} 
\newtheorem{definition}{Definition} 
\newtheorem{lemma}{Lemma} 
\newtheorem*{description*}{Problema} 
\newtheorem{remark}{Remark} 
\newtheorem{exercise}{Exercise} 
\newtheorem{assumption}{Assumption}


\begin{document}


\begin{theorem}[Initial approximation to the effect of the state graph size\footnote{Maybe this result can be improved (for a fixed explorative policy).}] 
    Let $\mathcal{X}$ be the set of nodes of the state graph underlying the generative process and
    let $N_{v}$ be the number of different nodes visited during the training of the GFlowNet\footnote{I suppose that
    this result can be improved; the result that I lay out here is just the best-case scenario. The problem is in how to
count the number of differently sampled paths on training.} . If this 
    training persists for $M$ epochs and at each epoch we visit at most $K$ different states, 
    then the probability that we visit at most a fraction $s$ of $\mathcal{X}$ satisfies 

    \begin{equation*}
        \mathbb{P}[N_{v} \ge s |\mathcal{X}|] \le \frac{MK}{s|\mathcal{X}|}. 
    \end{equation*}
\end{theorem}

\begin{proof}
    Just notice that $\mathbb{E}[N_{v}] \le MK$ and apply 
    Markov's inequality. 
\end{proof}

Empirically, I found that this probability is exponentially decaying on the size of the support. This is not surprising:
the training of the GFlowNet induces an open cluster within the state graph and it is well known from the field of
statistical mechanics that under the supercritical regime the probability of existing an open cluster is exponentially
decaying on the size of the cluster. 

\begin{theorem}[Effect of the size of state graph that are trees\footnote{Estou inspecionando o efeito do tamanho do grafo de estado no treinamento da GFlowNet. Estes são apenas rascunhos.}]
    Let $\mathbb{T}_{N}^{(K)}$ be a tree with depth $K$ and whose root has degree $N$ and all subsequent nodes have $N$ children each. Then, asymptotically as $K \rightarrow \infty$, the
    probability of visiting $H$ different states in $M$ iterations of an uniform policy decays exponentially in $H$, 

    \begin{equation*}
        \mathbb{P}[N_{v} \ge H] \le \exp\{-\alpha H\}, % . 
    \end{equation*}
    
    \noindent with $\alpha > 0$. Thus, the proper exploration of the distribution's support by a GFlowNet requires
    \textit{many} iterations. 
\end{theorem}

\begin{proof}
    Notice that, under an uniform exploratory policy, the probability of visiting a node at depth $d$ is
    $\frac{1}{N^{d}}$; this probability increases to $\frac{M}{N^{d}}$ for $M$ iterations of training. 
    For $d \ge \log \frac{M}{N} + 1$ and asymptotically as $K \rightarrow \infty$, the exploratory policy induces a
    percolation process under a subcritical regime and the number of explored nodes correspond to the size of the
    largest open cluster characterized by this process. It is well known that the probability of existing such a cluster
    decays exponentially on its size. Hence, the proportion of explored nodes within the state graph decays
    exponentially on the size of this graph.    
\end{proof}



In typical applications (as molecular generation), it is common to find
$|\mathcal{X}| \sim 10^{8}$ and $K = 20$; thus, the exploration of a large subset of the state graph would require
millions of iterations. 

\newpage

\section{Draft ideas: Eliezer}


\end{document}