\begin{abstract}
Image diffusion models have been adapted for real-world video super-resolution to tackle over-smoothing issues in GAN-based methods.
However, these models struggle to maintain temporal consistency, as they are trained on static images, limiting their ability to capture temporal dynamics effectively.
Integrating text-to-video (T2V) models into video super-resolution for improved temporal modeling is straightforward. 
However, two key challenges remain: artifacts introduced by complex degradations in real-world scenarios, and compromised fidelity due to the strong generative capacity of powerful T2V models (\textit{e.g.}, CogVideoX-5B). 
To enhance the spatio-temporal quality of restored videos, we introduce\textbf{~\name} (\textbf{S}patial-\textbf{T}emporal \textbf{A}ugmentation with T2V models for \textbf{R}eal-world video super-resolution), a novel approach that leverages T2V models for real-world video super-resolution, achieving realistic spatial details and robust temporal consistency.
Specifically, we introduce a Local Information Enhancement Module (LIEM) before the global attention block to enrich local details and mitigate degradation artifacts. 
Moreover, we propose a Dynamic Frequency (DF) Loss to reinforce fidelity, guiding the model to focus on different frequency components across diffusion steps. 
Extensive experiments demonstrate\textbf{~\name}~outperforms state-of-the-art methods on both synthetic and real-world datasets.\footnotemark
\footnotetext{$^*$Equal contributions. Work done during Rui Xie’s ByteDance internship. $\dagger$ indicates corresponding author.}
\end{abstract}
\vspace{-4mm}